Arrows for Parallel Computations

### 2017-11-27 paper/content/benchmarks.tex
minor

Because of the bad scaling behaviour for $2^{3217}-1$ beyond 64 threads, we also ran tests with input $2^{4253}-1$. Because of the long running time, we could do this only for the 128 and 256 threads. Therefore we do not show these results in Fig.~\ref{fig:bench-jacobi-dist} as we cannot properly compute a speedup without results for 1 thread. Nonetheless, for 128 threads, the benchmark took a mean of 9192.9\seconds, while with 256 threads, it only took a mean of 1649.1\seconds. This means that the 128 thread version ran more than 5.5 times slower than the 256 one, which suggests an IO limitation for this big input that is somewhat mitigated by adding more cores. However it still proves that PArrows continue to scale, even if not perfectly for this test program.
Comparing PArrows with Eden, for the larger input of $2^{4253}-1$, we see slightly bigger differences between Eden and PArrows: The differences for 128 and 256 coresare $-1005.18\seconds$ and $-94.33\seconds$ in favour of Eden, respectively. This maximum of $12.27\%$ slower PArrows runtime, however, was still in the (quite big) error bar of our measurements.

>>>>

Because of the bad scaling behaviour for $2^{3217}-1$ beyond 64 threads, we also ran tests with input $2^{4253}-1$. Because of the long running time, we could do this only for the 128 and 256 threads. Therefore we do not show these results in Fig.~\ref{fig:bench-jacobi-dist} as we cannot properly compute a speedup without results for 1 thread. Nonetheless, for 128 threads, the benchmark took a mean of 9192.9\seconds, while with 256 threads, it only took a mean of 1649.1\seconds. This means that the 128 thread version ran more than 5.5 times slower than the 256 one, which suggests an IO limitation for this big input that is somewhat mitigated by adding more cores. However it still proves that PArrows continue to scale, even if not perfectly for this test program.
Comparing PArrows with Eden, for the larger input of $2^{4253}-1$, we see slightly bigger differences between Eden and PArrows: The differences for 128 and 256 core sare $-1005.18\seconds$ and $-94.33\seconds$ in favour of Eden, respectively. This maximum of $12.27\%$ slower PArrows runtime, however, was still in the (quite big) error bar of our measurements.

### 2017-11-27 paper/content/topologySkeletons.tex
minor

This action is, however,  more complicated than in the ring case as we have one more dimension of inputs to be transformed. We first have to |shuffle| all the inputs to then pass it into |evalN conf (repeat (ptorus f))| which yields |[(d, fut a, fut b)]|. We can then unpack this shuffled list back to its original ordering by feeding it into the specific unshuffle arrow we created one step earlier with |arr length >>> arr unshuffle| with the use of |app :: arr (arr a b, a) c| from the |ArrowApply| type class. Finally, we unpack this matrix |[[[(d, fut a, fut b)]]| with |arr (map unzip3) >>> arr unzip3 >>> threetotwo| to get  |([[d]], ([[fut a]], [[fut b]]))|.

@@

ptorus :: (Arrow arr, Future fut a, Future fut b) =>
	arr (c, a, b) (d, a, b) -> arr (c, fut a, fut b) (d, fut a, fut b)
ptorus f = arr (\ ~(c, a, b) -> (c, get a, get b)) >>> f >>> arr (\ ~(d, a, b) -> (d, put a, put b))
\end{code} % $
\caption{Definition of the |torus| skeleton. The definitions of |lazyzip3|, |uncurry3| and |threetotwo| have been omitted and can be found in Fig.~\ref{fig:lazyzip3etc}}.
\label{fig:torus}
\end{figure}
As an example of using this skeleton \citep{Eden:SkeletonBookChapter02} showed the matrix multiplication using the Gentleman algorithm (\citeyear{Gentleman1978}). An adapted version can be found in Fig.~\ref{fig:torusMatMult}.
\begin{figure}[tb]
\begin{code}
type Matrix = [[Int]]

>>>>

This action is, however,  more complicated than in the ring case as we have one more dimension of inputs that needs to be transformed. In order to achieve that, we first have to |shuffle| all the inputs to then pass it into |evalN conf (repeat (ptorus f))| which yields |[(d, fut a, fut b)]|. We can then unpack this shuffled list back to its original ordering by feeding it into the specific unshuffle arrow we created one step earlier with |arr length >>> arr unshuffle| with the use of |app :: arr (arr a b, a) c| from the |ArrowApply| type class. Finally, we unpack this matrix |[[[(d, fut a, fut b)]]| with |arr (map unzip3) >>> arr unzip3 >>> threetotwo| to get  |([[d]], ([[fut a]], [[fut b]]))|.

@@

ptorus :: (Arrow arr, Future fut a, Future fut b) =>
	arr (c, a, b) (d, a, b) -> arr (c, fut a, fut b) (d, fut a, fut b)
ptorus f = arr (\ ~(c, a, b) -> (c, get a, get b)) >>> f >>> arr (\ ~(d, a, b) -> (d, put a, put b))
\end{code} % $
\caption{Definition of the |torus| skeleton. The definitions of |lazyzip3|, |uncurry3| and |threetotwo| have been omitted and can be found in Fig.~\ref{fig:lazyzip3etc}}.
\label{fig:torus}
\end{figure}
As an example of using this skeleton \citet{Eden:SkeletonBookChapter02} showed the matrix multiplication using the Gentleman algorithm (\citeyear{Gentleman1978}). An adapted version can be found in Fig.~\ref{fig:torusMatMult}.
\begin{figure}[tb]
\begin{code}
type Matrix = [[Int]]
